Theories of deliberated choice

### 2024-05-20 simple.tex
Kept single def indicativeness

\subsection{More properties}
A claim $C$ is \emph{informative} iff $\pi(C) = \set{\pi(\allleadstog) \suchthat {\allleadstog} \in C} ≠ (P \cup \set{\emptyset})^I$.
A claim $C$ is \emph{informative beyond counting} iff $\exists i \in I \suchthat \pi(C)(i) ≠ P \cup \set{\emptyset}$.
A claim $C$ is \emph{strongly informative beyond counting} iff $\exists i \in I \suchthat \pi(C)(i) \ssub P$.
% A claim $C$ is \emph{indicative} iff $\exists H \in \finite(\incr(\allhists)), i \in I \suchthat \pi(\set{{\allleadstog} \suchthat H \wsub \ileadstoinvg(P)} \cap C)(i) \ssub P$. Indicativity mandates that when knowing a finite number of observations about a given $i$, $C$ permits to deduce something (different than $\emptyset$) about $P_i$.
A claim $C$ is \emph{indicative} iff $\exists i \in I, \theta \in \Theta \cap (\allhists × \set{i} × P) \cap C^\wsub \suchthat \emptyset ≠ \pi(\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C)(i) \ssub P$. Indicativity mandates that when knowing a finite number of observations about a given $i$ compatible with any (possibly counter-factual) world in $C$, $C$ permits to deduce something (different than $\emptyset$) about $P_i$.
A claim $C$ is \emph{truly indicative} iff $\exists i \in I, \theta \in \Theta \cap (\allhists × \set{i} × P) \wsub {\allleadsto} \suchthat \emptyset ≠ \pi(\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C)(i) \ssub P$.
A claim $C$ is \emph{newly indicative} iff $\forall {\allleadstog} \in C: \exists i \in I, \theta \in \Theta \cap (\allhists × \set{i} × P) \wsub {\allleadstog} \suchthat \emptyset ≠ \pi(\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C)(i) \ssub P$. Indicativity mandates that when knowing a finite number of observations about a given $i$ compatible with any (possibly counter-factual) world in $C$, $C$ permits to deduce something (different than $\emptyset$) about $P_i$.

@@

  \begin{theorem}
    \label{th:suffIndic}
    The following two conditions on $\allleadsto$ are equivalent:
    \begin{enumerate}[label=T\ref{th:suffIndic}-{\arabic*}]
      \item \label{it:condHAI} for some $H \in \finite(\incr(\allhists))$ and $i_1 \in I \suchthat P_{i_1} \in P$, we have $\forall i_2 \suchthat P_{i_2} \neq P_{i_1}$ that $\exists \hist \in H \suchthat {\ileadsto[i_1]}(\hist) \neq {\ileadsto[i_2]}(\hist)$;
      \item \label{it:HAI} there exists a claim that holds and is anonymous and indicative.
    \end{enumerate}
  \end{theorem}
  \begin{proof}
    For \ref{it:condHAI} $\implies$ \ref{it:HAI},
    define $p = P_{i_1}$.
    Consider any claim $C$ such that $C^\text{HA} \wsub C \wsub C^\text{A}$ and $C$ is anonymous.
    % Define the claim $C$ as $\set{{\allleadstog} \in \allleadstos \suchthat \forall i \in I: {\ileadstog} \in \set{{\ileadsto[j]} \suchthat j \in I}}$ (or as anything included in it that includes $C^\text{HA}$ and is anonymous).
    The claim is indicative for $i = i_1$ and $\theta = \set{(\hist, i_1, {\ileadsto[i_1]}(\hist)) \suchthat \hist \in H}$ for the following reason.
    The attitudes in $C$ compatible with $\theta$ are $\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C \wsub \set{{\allleadstog} \suchthat {\ileadstog[i_1]} \in \set{{\ileadsto[j]} \suchthat P_j = p}}$, and $\pi(\set{{\allleadstog} \suchthat {\ileadstog[i_1]} \in \set{{\ileadsto[j]} \suchthat P_j = p}})(i_1) = p$.

@@

\subsection{Indicativeness}
\begin{theorem}
  Indicativeness does not imply true indicativeness, even when considering only anonymous and falsifiable claims that hold.
\end{theorem}
\begin{proof}
  Consider the situation where everybody is constant to $\phi$ except that emptyset to $¬\phi$.
  The claim “C HA or everybody is constant to $\phi$” is anonymous and falsifiable and indicative and holds, but is not truly indicative.
  Thus, the claim is “(everybody satisfies emptyset to $\phi$ or everybody satisfies emptyset to $¬\phi$) and (everybody satisfies every path but emptyset to $\phi$)”.
  It is falsifiable: any behavior function not in the claim satisfies “(somebody fails emptyset to $\phi$ and somebody fails emptyset to $¬\phi$) or (somebody fails every path but emptyset to $\phi$)”, which is observable in either case.
  It is indicative, as the behavior function “everybody constant to $\phi$” is what remains in the claim if assuming observing someone sending $\emptyset$ to $\phi$.

  But it is not truly indicative: this would require some truthful observation that leaves only a subset of the claim with no $\pi({\allleadstog})(i) = \emptyset$, which requires excluding “everybody satisfies emptyset to $¬\phi$”.
\end{proof}

Let us determine which situations admit anonymous and falsifiable claims that hold and are truly indicative (or other?).

@@

\begin{theorem}
  If an anonymous claim holds and is truly indicative, then $\exists i \in I \suchthat P_i ≠ \emptyset$.
\end{theorem}
\begin{proof}
  By anon and holding, we have $C^\text{HA} \wsub C$.
  By true indicativeness, for some $i$ and $\hist$ and $f \wsub {\ileadsto}$ on that $\intvl{\hist}$, all and some completion of $f$ in $C$ lead to $\pi({\allleadstog})(i) ≠ \emptyset$, thus, in particular, 
  ${\allleadsto}$ leads to $\pi({\allleadsto})(i) ≠ \emptyset$, 
  whence 
  $\exists i \in I \suchthat P_i ≠ \emptyset$.
\end{proof}

@@

\begin{conjecture}
  If $P_{i_1} = \phi$ with a single decisive argument $\ar \ileadstost \phi$ and $i_2$ satisfies ${\ileadsto[i_2]} = {\ileadsto[i_1]}$ except that $(\ar) \ileadsto[i_2] ¬\phi$, then there is no anonymous and truly indicative claim that holds.
\end{conjecture}
\begin{proof}
  True indicativeness requires that
  $\exists i \in I, \theta \in \Theta \cap (\allhists × \set{i} × P) \wsub {\allleadsto} \suchthat \emptyset ≠ \pi(\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C)(i) \ssub P$.
  If $\theta$ observes $(\ar_1, \ar)$
\end{proof}

>>>>

\subsection{More properties}
A claim $C$ is \emph{informative} iff $\pi(C) = \set{\pi(\allleadstog) \suchthat {\allleadstog} \in C} ≠ (P \cup \set{\emptyset})^I$.
A claim $C$ is \emph{informative beyond counting} iff $\exists i \in I \suchthat \pi(C)(i) ≠ P \cup \set{\emptyset}$.
A claim $C$ is \emph{strongly informative beyond counting} iff $\exists i \in I \suchthat \pi(C)(i) \ssub P$.
A claim $C$ is \emph{indicative} iff $\forall {\allleadstog} \in C: \exists i \in I, \theta \in \Theta \cap (\allhists × \set{i} × P) \wsub {\allleadstog} \suchthat \emptyset ≠ \pi(\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C)(i) \ssub P$. Indicativity mandates that for each (possibly counter-factual) world in $C$, when knowing a finite number of observations about a given $i$ compatible with, $C$ permits to deduce something (different than $\emptyset$) about $P_i$.

@@

  \begin{theorem}
    \label{th:suffIndic}
    The following two conditions on $\allleadsto$ are equivalent:
    \begin{enumerate}[label=T\ref{th:suffIndic}-{\arabic*}]
      \item \label{it:condHAI} for some $H \in \finite(\incr(\allhists))$ and $i_1 \in I \suchthat P_{i_1} \in P$, we have $\forall i_2 \suchthat P_{i_2} \neq P_{i_1}$ that $\exists \hist \in H \suchthat {\ileadsto[i_1]}(\hist) \neq {\ileadsto[i_2]}(\hist)$;
      \item \label{it:HAI} there exists a claim that holds and is anonymous and indicative.
    \end{enumerate}
  \end{theorem}
  \begin{proof}
    TODO check this, there is a problem with the definition and use of CA.

    For \ref{it:condHAI} $\implies$ \ref{it:HAI},
    define $p = P_{i_1}$.
    Consider any claim $C$ such that $C^\text{HA} \wsub C \wsub C^\text{A}$ and $C$ is anonymous.
    % Define the claim $C$ as $\set{{\allleadstog} \in \allleadstos \suchthat \forall i \in I: {\ileadstog} \in \set{{\ileadsto[j]} \suchthat j \in I}}$ (or as anything included in it that includes $C^\text{HA}$ and is anonymous).
    The claim is indicative for $i = i_1$ and $\theta = \set{(\hist, i_1, {\ileadsto[i_1]}(\hist)) \suchthat \hist \in H}$ for the following reason.
    The attitudes in $C$ compatible with $\theta$ are $\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C \wsub \set{{\allleadstog} \suchthat {\ileadstog[i_1]} \in \set{{\ileadsto[j]} \suchthat P_j = p}}$, and $\pi(\set{{\allleadstog} \suchthat {\ileadstog[i_1]} \in \set{{\ileadsto[j]} \suchthat P_j = p}})(i_1) = p$.

@@

\subsection{Indicativeness}
Let us determine which situations admit anonymous and falsifiable claims that hold and are indicative.

@@

@@

\begin{conjecture}
  If $P_{i_1} = \phi$ with a single decisive argument $\ar \ileadstost \phi$ and $i_2$ satisfies ${\ileadsto[i_2]} = {\ileadsto[i_1]}$ except that $(\ar) \ileadsto[i_2] ¬\phi$, then there is no anonymous and indicative claim that holds.
\end{conjecture}
\begin{proof}
  Indicativeness requires that
  $\exists i \in I, \theta \in \Theta \cap (\allhists × \set{i} × P) \wsub {\allleadsto} \suchthat \emptyset ≠ \pi(\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C)(i) \ssub P$.
  If $\theta$ observes $(\ar_1, \ar)$
\end{proof}
